---
title: "R Notebook"
output: html_notebook
---

2025.05.29 next_expected_data changed to next_expected_date but not saved.

```{r}
library(tidyverse)
```

## Load and preprocess

```{r}
drug =
  # "thiazide"
  # "b-blocker"
  # "ACEi"
  # "ARB"
  "CCB"
  # "ezetimibe"
  # "fibrate"
  # "statin"
  # "biguanide"
  # "sulfonylurea" #UNCHANGED
  # "thiazolidinedione"
  # "antidiabetic_drugs_excl_biguanide_sulfonylurea"
  # "insulin"
  # "bileacidsequestrant"
  # "nicotinicacid"
  # "omega3"

data = readr::read_tsv(paste0(drug, ".txt.gz"))
summary(data)
```

QC of issue_date

https://biobank.ndph.ox.ac.uk/showcase/exinfo.cgi?src=Data_providers_and_dates

N.B.
where clinical event or prescription date precedes participant date of birth it has been altered to 01/01/1901.
Where the date matches participant date of birth it has been altered to 02/02/1902.
Where the date follows participant date of birth but is in the year of their birth it has been altered to 03/03/1903.
Where the date was in the future this has been changed to 07/07/2037 as these are likely to have been entered as a place-holder or other system default.

```{r}
censoringdateleft  = as.Date("1938-01-01")
censoringdateright = as.Date("2017-08-31")
sum(is.na(data$issue_date))
sum(data$issue_date < censoringdateleft,  na.rm=TRUE)
sum(data$issue_date > censoringdateright, na.rm=TRUE)
data = data[
  !is.na(data$issue_date) &
    data$issue_date >= censoringdateleft &
    data$issue_date <= censoringdateright, ]
summary(data)
```

```{r}
table(!is.na(data$read_2), !is.na(data$bnf_code), !is.na(data$dmd_code))
```

```{r}
x = !is.na(data$read_2) & is.na(data$bnf_code) & is.na(data$dmd_code)
sum(x)
head(data[x, ])
```

```{r}
x = is.na(data$read_2) & !is.na(data$bnf_code) & is.na(data$dmd_code)
sum(x)
head(data[x, ])
```

```{r}
x = is.na(data$read_2) & is.na(data$bnf_code) & !is.na(data$dmd_code)
sum(x)
head(data[x, ])
```

```{r}
nrow(data)
length(unique(data$eid))
```

## Define unified drug code

```{r}
data$code =
  ifelse(
    !is.na(data$bnf_code),
    paste0("bnf_code ", data$bnf_code),
    ifelse(
      !is.na(data$read_2),
      paste0("read_2 ", data$read_2),
      paste0("dmd_code ", data$dmd_code)))
```

```{r}
x = data %>%
  group_by(eid, code, drug_name) %>%
  summarize(n = n(), .groups="drop")
nrow(x)
summary(x)
```

```{r}
y = x[x$n == max(x$n), ]
data[data$eid==y$eid[1] & data$code==y$code[1], ]
```

No obvious cutoff.

```{r}
plot(log10(sort(x$n)))
```

## Define issue_date_interval; time to next prescription

```{r}
data = data %>%
  group_by(eid, code, drug_name) %>%
  arrange(issue_date) %>%
  mutate(issue_date_interval = lead(issue_date) - issue_date)
```

```{r}
saveRDS(data, file=paste0(drug, ".rds"))
```

```{r}
data = readRDS(paste0(drug, ".rds"))
```

## Convert quantity (text) to total_units (numeric)

```{r}
source("calculate_total_units.R")
```

```{r}
x = sort(unique(data$quantity))
y = lapply(x, calculate_total_units)
foo = data.frame(quantity = x, total_units = unlist(y))
```

```{r}
data = left_join(x=data, y=foo)
```

```{r}
mean(! is.na(data$total_units)) / mean(! is.na(data$quantity))
mean(data$total_units <= 0, na.rm=TRUE)
data$total_units[data$total_units <= 0] = NA
summary(data$total_units)
```

```{r}
ggplot(
  data = data.frame(
    x = seq(0, 1, 0.001),
    y = quantile(data$total_units, seq(0, 1, 0.001), na.rm=TRUE))) +
  geom_hline(yintercept=c(7,28,56,84,112,168, 224,336), color='blue') +
  geom_point(aes(x = x, y = y), size=0.5) +
  scale_y_log10() +
  xlab("Quantile") +
  ylab("total_units (blue: 7,28,56,84,112,168)")
```

```{r}
if (drug %in% c("omega3")) {
  data$total_units[data$total_units > 600] = NA
} else {
  if (drug %in% c("bileacidsequestrant")) {
    data$total_units[data$total_units > 360] = NA
  } else {
    if (drug %in% c("nicotinicacid", "biguanide", "sulfonylurea")) {
      data$total_units[data$total_units > 336] = NA
    } else {
      data$total_units[data$total_units > 180] = NA
    }
  }
}

summary(data$total_units)
```

## Inspect issue_date_interval 

For insulin, 28 and 56 are frequent.

```{r}
sum(is.na(data$issue_date_interval)) # one per eid : code
sum(data$issue_date_interval==0, na.rm=TRUE)
summary(as.numeric(data$issue_date_interval, units="days"))
ggplot(
  data = data.frame(
    x = seq(0, 1, 0.001),
    y = quantile(as.numeric(data$issue_date_interval, units="days"), seq(0, 1, 0.001), na.rm=TRUE))) +
  geom_hline(yintercept=c(7,28,56,84), color='blue') +
  geom_point(aes(x = x, y = y), size=0.5) +
  scale_y_log10() +
  xlab("Quantile") +
  ylab("issue_date interval (blue: 7,28,56,84)")
```

## Plot issue_date_interval vs total_units; compute daily_units, next_expected_date

```{r}
ggplot(
  data=data,
  aes(x=as.numeric(issue_date_interval, units="days"), y=total_units)) +
  # geom_bin2d(bins=50) +
  stat_bin_2d(aes(fill = after_stat(log10(count))), bins=50) +
  scale_fill_viridis_c() +
  scale_x_log10() +
  scale_y_log10() +
  geom_abline(slope=1, intercept=0, color='red', alpha=0.75) +
  geom_abline(slope=1, intercept=log10(2), color='brown', alpha=0.75) +
  geom_abline(slope=1, intercept=log10(3), color='red', alpha=0.75) +
  geom_abline(slope=1, intercept=log10(4), color='red', alpha=0.75) +
  xlab("issue_date_interval [days]")
```

!!! TODO !!!
NEW daily_units

```{r}
# data$daily_units = ifelse(
#   data$total_units / as.numeric(data$issue_date_interval, units="days") > sqrt(2), 2, 1)

data$daily_units = ifelse(
  is.na(data$total_units),
  NA,
  ifelse(
    is.na(data$issue_date_interval),
    1,
    ifelse(
      (drug %in% c("bileacidsequestrant")) &
        data$issue_date_interval > 0 &
        data$total_units / as.numeric(data$issue_date_interval, units="days") > sqrt(6*5) &
        data$total_units %% 6 == 0,
      6,
      ifelse(
        (drug %in% c("bileacidsequestrant")) &
          data$issue_date_interval > 0 &
          data$total_units / as.numeric(data$issue_date_interval, units="days") > sqrt(5*4) &
          data$total_units %% 5 == 0,
        5,
        ifelse(
          (drug %in% c("bileacidsequestrant", "nicotinicacid", "omega3", "biguanide", "sulfonylurea")) &
            data$issue_date_interval > 0 &
            data$total_units / as.numeric(data$issue_date_interval, units="days") > sqrt(4*3) &
            data$total_units %% 4 == 0,
          4,
          ifelse(
            data$issue_date_interval > 0 &
              data$total_units / as.numeric(data$issue_date_interval, units="days") > sqrt(3*2) &
              data$total_units %% 3 == 0,
            3,
            ifelse(
              data$issue_date_interval > 0 &
                data$total_units / as.numeric(data$issue_date_interval, units="days") > sqrt(2) &
                data$total_units %% 2 == 0,
              2,
              1)))))))
  
data$next_expected_date = data$issue_date + data$total_units / data$daily_units
```

QUICK DIRTY SOLUTION FOR insulin

```{r}
if (drug=="insulin") {
  data$next_expected_date = data$issue_date + 84
}
```

```{r}
ggplot(
  data=data,
  aes(x=as.numeric(issue_date_interval, units="days"), y=total_units)) +
  geom_point(aes(color=factor(daily_units)), size=0.5, alpha=0.5) +
  scale_fill_viridis_c() +
  scale_x_log10() +
  scale_y_log10() +
  geom_abline(slope=1, intercept=0, color='red') +
  geom_abline(slope=1, intercept=log10(2), color='brown') +
  geom_abline(slope=1, intercept=log10(3), color='red') +
  geom_abline(slope=1, intercept=log10(4), color='red') +
  xlab("issue_date_interval [days]")
```

```{r}
x = data %>%
  filter(sum(!is.na(daily_units)) >= 5) %>%
  summarize(n=n(), sd = sd(daily_units, na.rm=TRUE), .groups='drop') %>%
  slice_max(sd, n=10)
x
```

```{r}
i=1
data[data$eid==x$eid[1] & data$code==x$code[1], ]
```

## Convert drug_name (text) to drug_name_brand (text) and drug_name_mg (numeric)

TODO: micrograms, mcg for stain
metformin 1 g modified-release tablets
g gram for bileacidsequestrant omega3

```{r}
source("parse_drug_name.R")
```

```{r}
x = sort(unique(data$drug_name))
y = do.call(rbind, lapply(x, parse_drug_name))
colnames(y) = c("drug_name_brand", "drug_name_mg")
foo = cbind(data.frame(drug_name = x), y)
summary(foo$drug_name_mg)
```

```{r}
data = left_join(x=data, y=foo)
```

```{r}
mean(! is.na(data$drug_name_mg))
table(! is.na(data$drug_name_mg), ! is.na(data$daily_units))
```

## Define daily_mg

```{r}
data$daily_mg = data$drug_name_mg * data$daily_units
mean(!is.na(data$daily_mg))
summary(data$daily_mg)
```

Below actually only chooses large SD w.r.t. daily_units, because drug_name_mg is constant in a group.

```{r}
x = data %>%
  filter(sum(!is.na(daily_mg)) >= 5) %>%
  summarize(n=n(), sd = sd(daily_mg, na.rm=TRUE), .groups='drop') %>%
  slice_max(sd, n=10)
x
```

```{r}
i=1
data[data$eid==x$eid[1] & data$code==x$code[1], ]
```


```{r}
saveRDS(data, file=paste0(drug, ".2.rds"))
```

```{r}
data = readRDS(paste0(drug, ".2.rds"))

data = rbind(
  readRDS("ACEi.2.rds"),
  readRDS("ARB.2.rds")
)
```

```{r}
min(data$total_units, na.rm=TRUE) > 0
max(data$total_units, na.rm=TRUE) <= 180
max(data$total_units, na.rm=TRUE) <= 336
```

## Combine consecutive prescriptions into a series (episode)

```{r}
data = data %>%
  ungroup() %>%
  drop_na(next_expected_date) %>%
  group_by(eid) %>%
  arrange(issue_date, next_expected_date) %>%
  # mutate(drug_free_preceding = issue_date - lag(next_expected_date))
  mutate(drug_free_preceding = as.numeric(issue_date) - cummax(lag(as.numeric(next_expected_date), default=-Inf))) #2025.06.27
```

Half (median) prescribed at exact renewal data.
This shows that our calculation of next_expected_date (and daily_units) is reasonable.
No clear cutoff for intermittence -> arbitrarily set to 14 days

```{r}
# sum(is.na(data$drug_free_preceding)) # one per eid
sum(data$drug_free_preceding==Inf) # one per eid
summary(data$drug_free_preceding)
ggplot(
  data = data.frame(
    x = seq(0, 1, 0.001),
    y = quantile(as.numeric(data$drug_free_preceding, units="days"), seq(0, 1, 0.001), na.rm=TRUE))) +
  geom_point(aes(x = x, y = y), size=0.5) +
  geom_hline(yintercept=c(7,14,28), color='blue') +
  ylim(-200, 200) +
  xlab("Quantile") +
  ylab("drug_free_preceding (blue: 7,14,28)")
```

```{r}
data
```

```{r}
data = data %>%
  mutate(new_flag = is.na(drug_free_preceding) | (drug_free_preceding > 14)) %>%
  mutate(scriptseries = cumsum(new_flag))
```

```{r}
scriptseries = data %>%
  group_by(eid, scriptseries) %>%
  summarize(eid=min(eid), start=min(issue_date), stop=max(next_expected_date), .groups="drop")
```

```{r}
scriptseries
```

```{r}
saveRDS(scriptseries, file=paste0("scriptseries.", drug, ".rds"))
```

